Master thesis

### 2017-06-09 docs/master-thesis.tex
Results: Disturbance & Appendix4

\subsubsection{Resilience}\label{resilience}

Tropical forest ecosystems provides numerous ecosystem services linked
to several ecosystem functions. We decided to describe simiulated
tropical forests in three major functions: forest dynamic, forest
production, and forest diversity. Forest dynamic was represented by
aboveground biomass (\(AGB\) in \(ton~C.ha^{-1}\)), basal area (\(BA\)
in \(m^2.ha^{-1}\)), total number of stem (\(N\)), number of stem above
\(10~cm\) diameter (\(N10\)), and number of stem above \(30~cm\)
diameter (\(N30\)). Forest production was represented by growth pimary
productivity (\(GPP\) in \(MgC.ha^{-1}\)), net pimary productivity
(\(NPP\) in \(MgC.ha^{-1}\)), tree autotrophic respiration in day
(\(Rday\) in \(MgC.ha^{-1}\)) and tree autotrophic respiration in night
(\(Rnight\) in \(MgC.ha^{-1}\)). Finally, forest diversity was
represented by \textbf{??}.

@@

The resilience of the system \(R(t)\) at the time \(t\) is described by
the ratio of recovery \(Recovery(t)\) at time \(t\) to loss suffered
\(Loss(t_d)\) at disturbance time \(t_d\). But in our peculiar case of
tropical forest ecosystems, the equilibrium used to calculate
\(Loss(t_d)\) can not be reduced to a specific time if the equilibrium
is dynamic. Consequently, to encompass undisturbed ecosystem variations
throught time, we simulated an undisturbed control ecosystem \(C\). And
the resilience of the system \(R(t)\) at the time \(t\) was defined as
the ratio of the ecosystem metric values in the disturbed simulation
\(X_T(t)\) over the the ecosystem metric values from the control
\(X_C(t)\) . Thus, the value of resilience \(R(t)\) is normalized for
all simulations and metrics. \(R(t)\) will be equal to \(R_{eq} = 1\)
when reaching the equilibrium value. Consequently we can calculate an
euclidean distance to equilibrium \(d_{eq}(t)\) as
\(d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}\).

@@

\(X_sp\) is the ecosystem variable value for one species either in
mixture \(X_{sp}(mixture)\) or in monoculture \(X_{sp}(monoculture)\).
\(P_{sp}\) is the proportion of the species in the mixture represented
by species relative abundance. Consequently, \(CE\) averages diversity
effects of all species presents in the mixture (both negatives and
positives). Whereas \(SE\) become positive when dominant species
outperform theimselves in mixture than in monoculture, and negative when
less dominant species outperform theimselves in mixture than in
monoculture \citep{Tobner2016}.

Recovery trajectories of ecosystem variable after disturbance were
partitioned between complementarity effect \(CE\) and selection effect
\(SE\). In order to do that, the design of experiment was repeated for
each species indivdually representing 652 simulations of monoculture.

@@

@@

>>>>

\subsubsection{Ecosystem functions}\label{ecosystem-functions}

Tropical forest ecosystems provides numerous ecosystem services linked
to several ecosystem functions. We decided to describe simiulated
tropical forests in two major functions: forest dynamic and forest
production. Forest dynamic was represented by aboveground biomass
(\(AGB\) in \(ton~C.ha^{-1}\)), basal area (\(BA\) in \(m^2.ha^{-1}\)),
total number of stem (\(N\)), number of stem above \(10~cm\) diameter
(\(N10\)), and number of stem above \(30~cm\) diameter (\(N30\)). Forest
production was represented by growth pimary productivity (\(GPP\) in
\(MgC.ha^{-1}\)), net pimary productivity (\(NPP\) in \(MgC.ha^{-1}\)),
tree autotrophic respiration in day (\(Rday\) in \(MgC.ha^{-1}\)) and
tree autotrophic respiration in night (\(Rnight\) in \(MgC.ha^{-1}\)).

@@

The resilience of the system \(R(t)\) at the time \(t\) is described by
the ratio of recovery \(Recovery(t)\) at time \(t\) to loss suffered
\(Loss(t_d)\) at disturbance time \(t_d\). But in our peculiar case of
tropical forest ecosystems, the equilibrium used to calculate
\(Loss(t_d)\) can not be reduced to a specific time if the equilibrium
is dynamic. Consequently, to encompass undisturbed ecosystem variations
throught time, we simulated an undisturbed control ecosystem \(C\). And
the resilience of the system \(R(t)\) at the time \(t\) was defined as
the ratio of the ecosystem metric values in the disturbed simulation
\(X_T(t)\) over the the ecosystem metric values from the control
\(X_C(t)\) . Thus, the value of resilience \(R(t)\) is normalized for
all simulations and metrics. \(R(t)\) will be equal to \(R_{eq} = 1\)
when reaching the equilibrium value. Consequently we can calculate an
euclidean distance to equilibrium \(d_{eq}(t)\) as
\(d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}\). Ecosystem euclidean distance
to equilibrium was calculated in a multi-dimensional space for the two
functions described above: forest dynamic (AGB, BA, N, N10, and N30) and
forest production (GPP, NPP, Rday, and Rnight). We then used cummulative
integral over time of eulcidean distance to equilibrium to assess
simulations resilience.

@@

\(X_{sp}\) is the ecosystem variable value for one species either in
mixture \(X_{sp}(mixture)\) or in monoculture \(X_{sp}(monoculture)\).
\(P_{sp}\) is the proportion of the species in the mixture represented
by species relative abundance. Consequently, \(CE\) averages diversity
effects of all species presents in the mixture (both negatives and
positives). Whereas \(SE\) become positive when dominant species
outperform theimselves in mixture than in monoculture, and negative when
less dominant species outperform theimselves in mixture than in
monoculture \citep{Tobner2016}. But similarly to resilience measurement,
biodiversity net effect \(NE\) can be in a dynamic equilibrium and vary
over time without disturbance. So in order to correctly assess selection
and complementarity effect in answer to disturbance, we normalized it by
undisturbed control ecosystem net effect \(NE_C\) to measure treatments
net effect resilience \(R(NE_T)\):

\begin{equation}
  R(NE_T) = \frac{NE_T}{NE_C} = \frac{SE_T}{NE_C} + \frac{CE_T}{NE_C}
  \label{eq:RNE}
\end{equation}

Resilience trajectories of ecosystem variable after disturbance were
partitioned between complementarity effect \(CE\) and selection effect
\(SE\). In order to do that, the design of experiment was repeated for
each species indivdually representing 652 simulations of monoculture.

@@

\subsubsection{Ecosystem functions}\label{ecosystem-functions-1}

We transformed all ecosystem outputs from the 240 disturbance
simulations in resilience metrics normalizing the treatment values by
their corresponding control (Figure \ref{fig:datatrans} A et B). We then
gathered ecosystem outputs by main ecosystem functions (forest dynamic,
forest production and forest diversity) to compute ecosystem distance to
equilibrium (Figure \ref{fig:datatrans} C). Finally, we integrated
distance to equilibrium in a cummulative sum over time (Figure
\ref{fig:datatrans} D).

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/datatrans-1.pdf}
\caption{\label{fig:datatrans}Ecosystem outputs data transformation.
Ecosystem outputs (\textbf{A}) are normalized by the control value over
time to calculate resilience (\textbf{B}); resilience of different
ecosystem outputs is then used in a multidimensional space to caluclate
ecosystem distance to equilibrium (\textbf{C}); finally distance ot
equilibirum is integrated over time in a cummulative sum (\textbf{D}).}
\end{figure}

The ranking was stable over time for the 240 simulations. So we used the
cummulative integral after 600 years \(Ieq_{600}\) as a measurement of
ecosystem resilience. We compared cummulative integral after 600 years
to communities taxonomic and functional diversity for each leavel of
disturbance (see Figure \ref{fig:gIeq}). We found that increased
functional diversity \citep[FDiv,][]{villeger_new_2008} was reducing
cummulative integral from ecosystem distance to forest dynamic
equilibrium after 600 years (\(Ieq_{600}\)). In addition, functional
eveness was complementary reducing \(Ieq_{600}\). Finally species
richness was not directly link to \(Ieq_{600}\). Effectivelly, low
species richness could result in variant \(Ieq_{600}\), but increased
species richness resulted in increased functional diversity and
consequently lower \(Ieq_{600}\). We found similar results for all
disturbance levels and other ecosystem functions (forest production and
forest diversity, see
\protect\hyperlink{appendix-4-disturbance-simulations}{Appendix 4:
Disturbance simulations}).

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/gIeq-1.pdf}
\caption{\label{fig:gIeq}Ecosystem resilience after 600 years with taxonomic
and functional diversity for different levels of disturbance.
Cummulative integral from ecosystem distance to forest dynamic
equilibrium after 600 years was represented against functional diversity
\citep[FDiv,][]{villeger_new_2008} for different level of disturbance
(25, 50 and 75\% of total basal area); dot shapes represents the species
richness whereas dot color represents functional eveness
\citep[FEve,][]{villeger_new_2008}.}
\end{figure}

\subsubsection{Biodiversity effect}\label{biodiversity-effect-1}

We measured all ecosystem outputs biodiversity net effect in disturbance
simulations by comparing theim to their species corresponding
monoculture simulations. The net effect was then partition between
selection and complementarity effect. We normalized treatments effect by
control net effect (see Table \ref{tab:tNE}) to measure their
aboveground biomass resilience to disturbance (see Figure
\ref{fig:gBE}). We found that complementarity effect was recovering net
effect in the first decades untill it disappeared after a century. On
the contrary selection effect was reduced or even removed by the
disturbance. It increased during the whole simulation and was greater
than complementarity effect only after decades. The time lag for which
complementarity effect was non null and greater than selection effect
was increasing with disturbance intensity. Finally 600 years after the
disturbance event biodiversity net effect was still not recovered for a
disturbance intensity greater than 25\% of basal area. We obtained those
results for aboveground biomass. We found similar results but with an
amlified signal for basal area (\(BA\)) and stem abundance (\(N\) but
with an inverted signal because of the forest self thinning) (see
\protect\hyperlink{appendix-4-disturbance-simulations}{Appendix 4:
Disturbance simulations}). Finally, forest growth primary productivity
(\(GPP\)) recovered in few years (proportionnaly to disturbance), and
its net effect was maintained by complementarity effect (see
\protect\hyperlink{appendix-4-disturbance-simulations}{Appendix 4:
Disturbance simulations}).

\begin{longtable}[]{@{}lrrll@{}}
\caption{\label{tab:tNE}Table 1: Biodiversity net effect mean value and
standard deviation for different ecosystem variable.}\tabularnewline
\toprule
variable & mean & standard deviation & name & unit\tabularnewline
\midrule
\endfirsthead
\toprule
variable & mean & standard deviation & name & unit\tabularnewline
\midrule
\endhead
agb & 32.721 & 17.839 & aboveground biomass &
\(tonC.ha^{-1}\)\tabularnewline
ba & 1.633 & 0.743 & basal area & \(m^2.ha^{-1}\)\tabularnewline
n & 103.880 & 231.623 & number of stems & \(n.ha^{-1}\)\tabularnewline
n10 & -6.815 & 13.003 & number of stems above 10cm dbh &
\(n.ha^{-1}\)\tabularnewline
gpp & 0.147 & 0.047 & growth primary production &
\(MgC.ha^{-1}\)\tabularnewline
npp & -0.041 & 0.036 & net primary production &
\(MgC.ha^{-1}\)\tabularnewline
Rday & 0.046 & 0.018 & autotrophic respiration during day &
\(MgC.ha^{-1}\)\tabularnewline
Rnight & 0.074 & 0.030 & autotrophic respiration during night &
\(MgC.ha^{-1}\)\tabularnewline
\bottomrule
\end{longtable}

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/gBE-1.pdf}
\caption{\label{fig:gBE}Resilience of complementarity and selection effects.
Complementarity effect (CE) and selection effect (SE) where normalized
by control net effect (NEc), thus measuring their resilience over time.}
\end{figure}

@@

\hypertarget{appendix-4-disturbance-simulations}{\section{Appendix 4:
Disturbance simulations}\label{appendix-4-disturbance-simulations}}

\subsection{Ecosystem functions}\label{ecosystem-functions-2}

Appendix 4 presents ecosystem resilience after 600 years with taxonomic
and functional diversity for different levels of disturbance. It
encompass all functionnal diversity components \citep[FRIC, FEve, FDiv,
and FDis,][]{villeger_new_2008}. And it presents results for both forest
dynamic (Figure \ref{fig:A4IeqStructureAll}) and forest production
(Figure \ref{fig:A4IeqProductionAll}).

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A4IeqStructureAll-1.pdf}
\caption{\label{fig:A4IeqStructureAll}Figure S1.1: Ecosystem resilience
after 600 years with taxonomic and functional diversity for different
levels of disturbance. Cummulative integral from ecosystem distance to
forest dynamic equilibrium after 600 years was represented against
functional diversity \citep[FRIC, FEve, FDiv, and
FDis,][]{villeger_new_2008} for different level of disturbance (25, 50
and 75\% of total basal area); dot shapes represents the species
richness.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A4IeqProductionAll-1.pdf}
\caption{\label{fig:A4IeqProductionAll}Figure S1.2: Ecosystem resilience
after 600 years with taxonomic and functional diversity for different
levels of disturbance. Cummulative integral from ecosystem distance to
forest production equilibrium after 600 years was represented against
functional diversity \citep[FRIC, FEve, FDiv, and
FDis,][]{villeger_new_2008} for different level of disturbance (25, 50
and 75\% of total basal area); dot shapes represents the species
richness.}
\end{figure}

\subsection{Biodiversity effect}\label{biodiversity-effect-2}

Figure \ref{fig:A4allBE} presents the resilience of complementarity and
selection effects for different ecosystem metrics (AGB, BA, N, GPP and
NPP).

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A4allBE-1.pdf}
\caption{\label{fig:A4allBE}Resilience of complementarity and selection
effects. Complementarity effect (CE) and selection effect (SE) where
normalized by control net effect (NEc), thus measuring their resilience
over time for different ecosystem variables (AGB, BA, N, GPP).}
\end{figure}
